Debugging of RxJS-based Applications

### 2020-09-29 paper.tex
Remove Percentages

RP is a declarative programming paradigm that has its origins in FP. While engineers use imperative programming languages to specify every step \emph{how} a program has to do something, declarative languages allow to describe \emph{what} the program should achieve ultimately. A runtime system then figures out a way to satisfy that description and executes it. RP functionality is usually provided in form of a language extension for a specific programming language (e.g. REScala for Scala\cite{10.1145/2577080.2577083}) or as a library (e.g. RxJS for JavaScript\cite{rxjs})

@@

The survey responses available in Table~\ref{tab:subject-survey} showed that 75\% (S2, S3 and S4) of the subject population had two or more years of experience with RxJS. Where all of them use RxJS to develop frontend applications, S4 declared having used  RxJS for backend development as well. When asked what tools they usually use for debugging, S2, S3 and S4 stated to use the traditional debugger of their IDE. S1 and S3 leverage additional tracing functionality of rxjs-spy, and all four of our subjects use manual log statements.

@@

We were able to observe how all subjects predominantly used manual source code augmentation by adding trace logs. Only 50\% of them used traditional debugging utilities in order to inspect the program's state at runtime. All subjects used the new information gained to refine their hypothesis about underlying problems before they started a new iteration in the debugging process. We could not observe the extraction to and reintegration from an external tool. All subjects exhibited the debugging behavior described in our hypothesis.

@@

More than 50\% of our peers throughout the interviews, war story reports and the experiments after-action survey stated to know about specific RxJS RP debugging tools. It was apparent that all subjects during the observational study refrained from using any of them, though. It is our speculation that the subjects knowing about specific tools held themselves back from using them because they perceived the effort of setting them up (e.g., installing and configuring rxjs-spy) as too time-consuming. Not having the ``right'' tool available without significant additional effort is also what we interpret from the statement by S4: Though they would have started to extract parts of the data-flow and inspect it with other tools, they would have done so only after the 25 minutes of the block expired; Hence a more accessible way allowing such analyses would have influenced the behavior of the subject.

@@

The observation that 50\% of our study subjects tried to debug an RP application with traditional, imperative-centric debugger utilities, as well as related statements from the interviews and war stories, strengthened our assumption regarding tool integration. Engineers expect the debugging tools they know and rely on to give correct insight on every program, no matter the paradigm (imperative or declarative) with which it got implemented.

@@

In this paper, we have explored how software engineers debug data-flow-oriented programs implemented using RxJS. We presented an observational study to validate a hypothesis based on the outcome of ten individual interviews and hands-on experience reports from software engineering professionals. More than 50\% of the engineers we worked with during our research told us that they know of the existence of specific debugging tools for RP with RxJS. Nonetheless, the experiment conducted with four participants allowed us to prove that engineers augment source code manually with trace logs instead of using such specialized utilities.

>>>>

RP is a declarative programming paradigm that is strongly influenced by FP. While engineers use imperative programming languages to specify every step \emph{how} a program has to do something, declarative languages allow to describe \emph{what} the program should achieve ultimately. A runtime system then figures out a way to satisfy that description and executes it. RP functionality is usually provided in form of a language extension for a specific programming language (e.g. REScala for Scala\cite{10.1145/2577080.2577083}) or as a library (e.g. RxJS for JavaScript\cite{rxjs})

@@

The survey responses available in Table~\ref{tab:subject-survey} showed that S2, S3 and S4 had two or more years of experience with RxJS. Where all of them use RxJS to develop frontend applications, S4 declared having used  RxJS for backend development as well. When asked what tools they usually use for debugging, S2, S3 and S4 stated to use the traditional debugger of their IDE. S1 and S3 leverage additional tracing functionality of rxjs-spy, and all four of our subjects use manual log statements.

@@

We were able to observe how all subjects predominantly used manual source code augmentation by adding trace logs. Two of them used traditional debugging utilities in order to inspect the program's state at runtime. All subjects used the new information gained to refine their hypothesis about underlying problems before they started a new iteration in the debugging process. We could not observe the extraction to and reintegration from an external tool. All subjects exhibited the debugging behavior described in our hypothesis.

@@

More than 50\% of our 14 peers throughout the interviews, war story reports and the experiments after-action survey stated to know about specific RxJS RP debugging tools. It was apparent that all subjects during the observational study refrained from using any of them, though. It is our speculation that the subjects knowing about specific tools held themselves back from using them because they perceived the effort of setting them up (e.g., installing and configuring rxjs-spy) as too time-consuming. Not having the ``right'' tool available without significant additional effort is also what we interpret from the statement by S4: Though they would have started to extract parts of the data-flow and inspect it with other tools, they would have done so only after the 25 minutes of the block expired; Hence a more accessible way allowing such analyses would have influenced the behavior of the subject.

@@

The observation that two out of four of our study subjects tried to debug an RP application with traditional, imperative-centric debugger utilities, as well as related statements from the interviews and war stories, strengthened our assumption regarding tool integration. Engineers expect the debugging tools they know and rely on to give correct insight on every program, no matter the paradigm (imperative or declarative) with which it got implemented.

@@

In this paper, we have explored how software engineers debug data-flow-oriented programs implemented using RxJS. We presented an observational study to validate a hypothesis based on the outcome of ten individual interviews and hands-on experience reports from software engineering professionals. More than 50\% of the 14 engineers we worked with during our research told us that they know of the existence of specific debugging tools for RP with RxJS. Nonetheless, the experiment conducted with four participants allowed us to prove that engineers augment source code manually with trace logs instead of using such specialized utilities.
